Write-and-f-array: implementation and an application

### 2013-08-31 msc.tex
listing moved & footnote->main text

Here we differentiate between threads: the index p in \fn{inc} corresponds to the current thread's ID from range
$\{0, \ldots, P-1\}$. Let's see first that this implementation actually serializes to the increment-only counter.
Obviously, we need to choose the call to WRITE as the serialization point of \fn{inc}. From the monotonicity
of entries of T we can infer that \fn{read} returns a value between (inclusive) the number of calls to \fn{inc} that have
serialized before \fn{read} has started and finished. Thus if a call to \fn{read} returns x, there is a point during its
execution when exactly x \fn{inc}s had been serialized. We can choose this point to be the serialization point of
\fn{read}\footnote{Note that it need not correspond to an action executed by the implementation of \fn{read}}.

@@

\begin{Listing}
\begin{algorithmic}[1]
\State\var{V}{\int}
\Function{fetch-and-add}{$x$}
	\State $r \gets V$
	\State $V \gets V + x$
	\State\Return $r$
\EndFunction
\end{algorithmic}
\caption{Specification of semantics of fetch-and-add}
\label{faa-spec}
\end{Listing}

We then use this object to implement a fetch-and-add object. Fetch-and-add is a generalization of the increment-only counter we've seen. Its semantics are specified in Listing~\ref{faa-spec}.
Fetch-and-add objects can be used to produce unique identifiers, implement
mutual exclusion, barrier synchronization\cite{faa-sync}, or work queues\cite{faa-queue}. Its known wait-free shared memory implementations for $P$ processes have $\Omega(P^2)$ memory complexity and $O(\log^2 P)$ step complexity
of the operation \cite{ellen-fai}\cite{closed-object}. They also need to employ complicated memory management techniques to bound their memory use. Our implementation reduces the memory complexity to $O(P\log P)$ while maintaining polylogarithmic step complexity.

>>>>

Here we differentiate between threads: the index p in \fn{inc} corresponds to the current thread's ID from range
$\{0, \ldots, P-1\}$. Let's see first that this implementation actually serializes to the increment-only counter.
Obviously, we need to choose the call to WRITE as the serialization point of \fn{inc}. From the monotonicity
of entries of T we can infer that \fn{read} returns a value between (inclusive) the number of calls to \fn{inc} that have
serialized before \fn{read} has started and finished. Thus if a call to \fn{read} returns x, there is a point during its
execution when exactly x \fn{inc}s had been serialized. We can choose this point to be the serialization point of
\fn{read}. Note that it need not correspond to an action executed by the implementation of \fn{read}.

@@

We then use this object to implement a fetch-and-add object. Fetch-and-add is a generalization of the increment-only counter we've seen. Its semantics are specified by the listing below:

%\begin{Listing}
\begin{algorithmic}[1]
\State\var{V}{\int}
\Function{fetch-and-add}{$x$}
	\State $r \gets V$
	\State $V \gets V + x$
	\State\Return $r$
\EndFunction
\end{algorithmic}
%\caption{Specification of semantics of fetch-and-add}
%\label{faa-spec}
%\end{Listing}

Fetch-and-add objects can be used to produce unique identifiers, implement
mutual exclusion, barrier synchronization\cite{faa-sync}, or work queues\cite{faa-queue}. Its known wait-free shared memory implementations for $P$ processes have $\Omega(P^2)$ memory complexity and $O(\log^2 P)$ step complexity
of the operation \cite{ellen-fai}\cite{closed-object}. They also need to employ complicated memory management techniques to bound their memory use. Our implementation reduces the memory complexity to $O(P\log P)$ while maintaining polylogarithmic step complexity.
